When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America

### 2020-08-12 code/Submissionv2.Rmd
commit

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
```
\newpage

@@

```{r summary, include = FALSE}
# Read in the Data
dataset <- read_csv(here("data", "MilEditorials24June.csv"))
dataset %>% print()
# Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
adjdataset <- dataset %>%
  mutate_at(
    .vars = vars(War_Ops:Fig_Nme),
    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
) %>% mutate(
    PubYr = lubridate::year(PubDate)
) %>% view()

@@

@@

```{r second level analysis, include = FALSE}

# Present Results of Result and Justification
secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
  select (Author, Title, Source, PubYr, Result, Justification) %>% 
  view()
```

@@

The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years. Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how military elites communicate with the American public. Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did. Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency. These trends are captured in Figure 4, below, which shows the number (left) and proportion (right) of opinion pieces authored by retired military officers over the past 40 years that addressed issues of social policy, domestic policy, or specific political figures or administrations. The graphs tell an interesting story: the number and proportion were both basically 0 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011. After that, both the number and proportion have risen sharply. Nearly 40% of all opinion pieces in 2019 published in major US newspapers by retired military officers were written about topics that fell outside of traditional national security topics.

>>>>

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
```
\newpage

@@

```{r summary, include = FALSE}
# Read in the Data
dataset <- read_csv(here("data", "MilEditorials10Aug.csv"))
dataset %>% print()
# Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
adjdataset <- dataset %>%
  mutate_at(
    .vars = vars(War_Ops:Fig_Nme),
    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
) %>% mutate(
    PubYr = lubridate::year(PubDate)
) %>% view()

@@

#adjust the data to account for polarization by creating a new dataset called polaradjdataset
#create variables that capture ideological distance in each the House and Senate along
#each of the two dimensions according to Poole and Rosenthal's data
polaradjdataset <-adjdataset %>% mutate (
  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1),
  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1),
  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2),
  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)
) %>% view()

@@

```{r second level analysis, include = FALSE}
# Present Results of Result and Justification
secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
  select (Author, Title, Source, PubYr, Ques_Prof, Justification) %>% 
  view()

#Now Prep the data that I need to analyze polarization levels.  I will try first by 
# eliminating the NA values but I will do this by indexing the first 334 rows of the 
# data, which corresponds to the number of observations that I have.  I do this by #creating a new data set called slimdata

slimdata <- polaradjdataset[1:334,] %>% 
  view()

#next, conduct a logit regression where the DV is values of Ques_Prof along
#several IVs

logmodel <- glm(
  Ques_Prof ~ Inst_Cred_Gallup + House_distance_dim1 + MilPSoc + Dom_P + Fig_Nme, 
  family=binomial(link="logit"), 
  data=slimdata
  )

tidy(logmodel)
glance(logmodel)

# now  make prediction data using the mean institutional credibility scores.  Vary #polarization and the three topics I am interested in.

predictiondata <- tibble(House_polarization_level=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Credibility = mean(slimdata$Inst_Cred_Gallup)) %>%
crossing(Ques_Prof = c(0,1), MilPSoc = c(0,1), Dom_P = c(0,1), Fig_Nme = c(0,1))

predictiondata

# actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec

augment(logmodel, newdata = predictiondata)

predictions <- augment(logmodel, newdata=predictiondata) %>% 
  mutate(
    .prob=plogis(.fitted),
    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
    pred_final_rec = case_when(
      .prob >= .5 ~ "Likely Violation of Professionalism",
      .prob < .5 ~ "Violation of Professionalism not likely"
    )
  ) 
predictions %>% arrange(.prob, .fitted, Report, Night, desc(.fitted)) %>% 
  view()

@@

The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years. Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how military elites communicate with the American public. Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did. Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency. These trends are captured in Figure 4, below, which shows the number (left) and proportion (right) of opinion pieces authored by retired military officers over the past 40 years that addressed issues of social policy, domestic policy, or specific political figures or administrations. The graphs tell an interesting story: the number and proportion were both basically zero 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011. After that, both the number and proportion have risen sharply. Nearly 40% of all opinion pieces in 2019 published in major US newspapers by retired military officers were written about topics that fell outside of traditional national security topics.
